Manuscript Figures

### 2024-04-05 manuscript_figures.Rmd
updates

@@

## A- Unsupervised clustering of Proteomics and Metabolomic data combined

```{r,fig.height = 8, fig.width = 12}
##Proteomics data ##
pl = read_csv("01_r_processed_data/quant_msstats_protein_level_data.csv") %>% 
  select(tmt_id = sample_id,everything())

clean_clinical_wide = clean_clinical %>% 
  select(sample_id,death_during_admission,Gene,Abundance) %>% 
 tidyr::pivot_wider(names_from = Gene,
              values_from = Abundance) %>% 
  #removing the samples without death during admission info
  filter(!is.na(death_during_admission)) %>% 
  #filter(sample_id.y != "S49",
         #sample_id.y != "S76") %>% 
  dplyr::select_if(~ !any(is.na(.))) %>% 
  #needs to be a dataframe or EFS freaks out. 
  as.data.frame()

### Combining all data ### 
all= clean_clinical_wide

# Getting into a format where we can do the clustering ## 
clust_data = all[,3:283] %>% 
   #Remove columns that only contain 0s 
  select_if(colSums(.) != 0) %>% 
  as.data.frame()

## Fixing the row names so they match for the heatmap ##
rownames(annotation) = all$sample_id

rownames(clust_data) = all$sample_id

#calculating distance matrix
d = dist(clust_data)
## Generating a dendrogram without a heatmap ##
library(dendextend)

# Plotting for manuscript
pdf("02_figures/f6_prediction_of_clinical_outcome_proteomics/f6a.pdf",height = 8,
    width = 12)
plot(dend)
dend %>% rect.dendrogram(k=2, 
                           border = 8, lty = 5, lwd = 2)
dev.off()

#Plotting for this document
plot(dend)
dend %>% rect.dendrogram(k=2, 
                           border = 8, lty = 5, lwd = 2)

# Checking stats to see if there are significant differences between clusters when k=2
xtab <- as.table(rbind(c(4, 8), c(13, 53)))
dimnames(xtab) <- list(
  group = c("k1", "k2"),
  death = c("yes", "no")
)

## B Volcano plot of Proteomics

@@

## D - GO Enrichment in Live

@@

#Have trouble manipulating the pdf in inkscape
save_pheatmap_pdf(p,"03_supplemental/s9_sequencing_based_prediction_of_clinical_outcome/s9a.pdf")
save_pheatmap_png(p,"03_supplemental/s9_sequencing_based_prediction_of_clinical_outcome/s9a.png")

@@

save_pheatmap_pdf(p2,"03_supplemental/s9_sequencing_based_prediction_of_clinical_outcome/s9b.pdf")
save_pheatmap_png(p2,"03_supplemental/s9_sequencing_based_prediction_of_clinical_outcome/s9b.png")
```

@@

>>>>

## ? Are proteins involved in the gene ontology results confounded?

```{r,fig.height = 8, fig.width = 8}
# Extracting the GO terms in 
enriched_in_faecium = enrich@result %>% 
  filter(p.adjust <= 0.05)

#Looking at the proteins in the cholesterol portions of the network
proteins_in_network = c("APOC3","APOE","APOC1","APOC2","PCYOX1","CLU")

stat = metadata_assesment(md_assesment_data,
                          metric = Gene,
                          sample_id = sample_id,
                          features = proteins_in_network,
                          outcome = Abundance,
                          outcome_c = "Abundance")

## Plotting ###
t = stat_final %>% 
  ggplot(aes(reorder(variable,p.adj),-log10(p.adj)))+
  geom_col()+
  geom_hline(yintercept = -log10(0.05),linetype = "dashed", color = "red")+
  coord_flip()+
  facet_wrap(~Gene,nrow = 2) +
  xlab("metadata_field")

t

for(i in proteins_in_network){

  f = filter(md_assesment_data,Gene %in% i,
             pathogen %in% c("faecalis","faecium"),
             transplant_type == "none")

stat = f %>% 
  rstatix::t_test(Abundance ~ pathogen,) %>% 
  #rstatix::add_significance() %>% 
  rstatix::add_xy_position(fun = "median")

p1 = ggplot(f,aes(pathogen,Abundance)) + 
  geom_violin()+
  geom_point()+
  ggtitle(i)+
  ggpubr::stat_pvalue_manual(stat,)

t2 = do.call("grid.arrange", plotlist,)
```

##? Are differentially abundant proteins the same ones that are different when comparing healthy to infected?

healthy_v_infected = readr::read_csv("01_r_processed_data/Infected_vs_Healthy.csv")

up_in_infected = healthy_v_infected %>% 
  filter(adj.pvalue <= 0.05,
         log2FC>0) %>% 
  pull(uniprotid)

down_in_infected =  healthy_v_infected %>% 
  filter(adj.pvalue <= 0.05,
         log2FC<0) %>% 
  pull(uniprotid)

faecalis_v_faecium = readr::read_csv("01_r_processed_data/faecalis_v_faecium.csv") %>% 
  left_join(protein_to_gene,by = c("uniprotid" = "ProteinID")) %>% 
  mutate(also_in = case_when(uniprotid %in% up_in_infected ~ "infected",
                             uniprotid %in% down_in_infected ~ "healthy",
                             TRUE ~ "distinct"))

  vp = faecalis_v_faecium %>% 
  ggplot(aes(log2FC,-log10(adj.pvalue),color = also_in))+
  geom_point(size = 0.5)+
  geom_hline(yintercept = -log10(0.05),linetype = "dashed",color = "red")+
  xlab("Log2FC Faecalis / Faecium")+
  ggprism::theme_prism(base_size = 10)

vp

@@

## A Volcano plot of Proteomics

@@

## C - GO Enrichment in Live

@@

#Have trouble manipulating the pdf in inkscape
save_pheatmap_pdf(p,"03_supplemental/s9_sequencing_based_prediction_of_clinical_outcome/s9a.pdf")
#save_pheatmap_png(p,"03_supplemental/s9_sequencing_based_prediction_of_clinical_outcome/s9a.png")

@@

save_pheatmap_pdf(p2,"03_supplemental/s9_sequencing_based_prediction_of_clinical_outcome/s9b.pdf")
#save_pheatmap_png(p2,"03_supplemental/s9_sequencing_based_prediction_of_clinical_outcome/s9b.png")
```

@@

####C- Remaining Transplant Type associations

```{r,fig.height =8, fig.width = 8}

transplant_type_associated = c("APOC3","AZGP1","PCYOX1","RBP4","SERPINC1")

sum_annotated = left_join(protein_sum,anot,by = c("uniprotid" = "Entry")) %>% 
  left_join(md,by = c("BioReplicate" = "sample_id")) %>% 
  filter(Gene %in% transplant_type_associated,
         transplant_type == "none")

for(i in transplant_type_associated){

stat = f %>% 
  rstatix::t_test(Abundance ~ condition,) %>% 
  #rstatix::add_significance() %>% 
  rstatix::add_xy_position(fun = "median")

p1 = ggplot(f,aes(condition,Abundance)) + 
  geom_violin()+
  geom_point()+
  ggtitle(i)+
  ggpubr::stat_pvalue_manual(stat,)

s10c = do.call("grid.arrange", plotlist,)

ggsave("03_supplemental/s10_checking_confounding_variables/s10c.pdf",s10c)
```

#### D- Metabolite Associations

```{r,fig.height =8, fig.width = 8}
metabolomics_data = read_csv("01_r_processed_data/normalized_metabolomics.csv")

m = inner_join(metabolomics_data,gnps_annotations,by = "row_id") %>% 
  filter(condition != "Healthy",
         transplant_type == "none")

transplant_type_associated = c("RID_21931","RID_22049","RID_22413","RID_22552","RID_11530","RID_26935")

for(i in transplant_type_associated){

  f = filter(m,row_id %in% i)

  name= unique(f$compound_name_cleaned)

stat = f %>% 
  rstatix::t_test(norm_value ~ condition) %>% 
  rstatix::add_significance() %>% 
  rstatix::add_xy_position()

p1 = ggplot(f,aes(condition,norm_value)) + 
  geom_violin()+
  geom_point()+
  ggtitle(paste0(name))+
  ggpubr::stat_pvalue_manual(stat)

s10d = do.call("grid.arrange", plotlist,)

ggsave("03_supplemental/s10_checking_confounding_variables/s10d.pdf",s10d)

m

```{r}
md = read_csv("00_r_input_data/clinical_metadata.csv") %>% 
  filter(transplant_type == "none",
        condition != "Healthy") %>% 
  select(sample_id,condition,transplant_type)
md
```
